Do we need detailed demographic data to forecast plant population responses to climate change?

### 2014-12-19 manuscript/manuscript.Rmd
writing methods

### Stastical models of vital rates
The first step in building our single-species population models was to fit statistical models of vital rates at both levels of inference: individual-level and quadrat-level. Here we describe the general statistical approach at the individual-level, but the same approach applies at the quadrat-level. We modeled survival and growth as functions of genet size (quadrat cover in the quadrat-level models) and climate covariates (described in more detail below). We maintained a consistent random effects structure for both models that included three terms: (1) a random year effect on the intercept, (2) a random year effect on the coefficient for plant cover (either individual or quadrat level), and (3) a random effect of group (see Data set description) on the intercept. Vital rates for each species are modeled separately.

We used logistic regression to model survival probability:

\begin{equation}
logit(s) = \beta_{0,t} + \beta_{s,t}x + \beta_{Q} + \beta_{c,1}\theta_{1,t}\\
+ \cdots +  \beta_{c,i}\theta_{i,t} + \varepsilon_{t}
\end{equation}

where $x$ is the log of genet size (quadrat areal cover), $\beta_{0,t}$ is a year-specific intercept, $\beta_{s,t}$ is the year-specific slope parameter for size, $\beta_{Q}$ is the random effect of quadrat group location, $\theta$ is a matrix of $i$ climate effects over $t$ years, $\beta_{c,i}$ is the fixed parameter for the effect of the $i$th climate covariate, and $\varepsilon_{t}$ is the error term. At the quadrat level we did not fit random year effects for survival because those parameters did not reach convergence (see "Computing").

We modeled growth as gaussian process describing genet size (or quadrat cover) at time $t+1$ as a function of size at $t$ and climate covariates:

\begin{equation}
x_{t+1} = \beta_{0,t} + \beta_{s,t}x_{t} + \beta_{Q} + \beta_{c,1}\theta_{1,t}\\
+ \cdots +  \beta_{c,i}\theta_{i,t} + \varepsilon_{t}
\end{equation}

where $x$ is genet size and all other paramters are as described for the survival regression. For the quadrat-level approach we modeled growth as a process describing proportional cover within a quadrat at time $t+1$ as a function of proportional cover at time $t$ and climate covariates. Thus, instead of a gaussian process, with a normal likelihood, we modeled growth at the quadrat level as above but with a beta likelihood and a logit link to the linear predictors.

While our approach for modeling survival and growth is similar at the individual and quadrat levels, the addition of new genets at the inidividual level or occupied sites at the quadrat level requires two separate approaches.

### Including climate covariates (and avoiding model selection)
For both the individual-level IPM (IPM) and the quadrat-based IBM (QBM) we included four climate covariates in each vital rate regression: fall through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1 and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is the observation year. We included climate covariates as additive effects. We did not consider interactions among climate covariates or among climate covariates and plant size or proportional cover. Climate effects ($\beta_{c}$s) were modeled hierarchically in each vital rate regression so that each species specific climate effect is drawn from an interspecific climate effect distribution. This was necessary to achieve convergence at the quadrat level and we used the same approach at the individual level for uniformity.

### Model fitting
All of our analyses (model fitting and simulating) were conducted in Program R. We used the MCMC sampler in JAGS to estimate the posterior distributions of model parameters. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 50,000 iterations, after a 50,000 iteration burn in. We assessed convergence visually and using the Gelman diagnostic in the R package `coda.' Scale reduction factors for all parameters were less than 1.02, indicating convergence. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each chain for all parameters to be used as randomly drawn values during population simulation.

@@

----------

>>>>

In this paper, we model populations based on two levels of data: individual and quadrat. The individual data is the "raw" data. For the quadrat level we data we simply sum individual areal cover for each quadrat by species. This is equivalent to a perfect census of quadrat percent cover, so we do not need to consider measurement error. Based on these two datasets we can compare population models built using individual level data and aggregated quadrat level data.

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual level data we fit models for survival, growth, and recruitment of new individuals for each species. At the quadrat level we fit analagous models of extinction probability, percent cover increase/decrease, and quadrat colonization for each species. We describe the statistical models seprately since fitting the models required different approaches at the individual and quadrat levels. All models contain four climate covariate that we chose \emph{a priori}: fall through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1 and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is the observation year

#### Vital rate models: individual level
We used logistic regression to model survival probability ($S$) of genet $i$ from species $j$ in quadrat group $Q$ from time $t$ to $t+1$:

\begin{equation}
\text{logit}(S_{ijQ,t}) = \gamma^{S}_{j,t} + \phi^{S}_{jQ} + \beta^{S}_{j,t}x_{ij,t} + \omega^{S}_{j}w_{ij,t} + \theta^{S}_{jk}C_{k,t} + \varepsilon^{S}_{t}
\end{equation}

where $x_{ij,t}$ is the log of genet size, $\gamma^{S}_{j,t}$ is a year-specific intercept, $\beta^{S}_{j,t}$ is the year-specific slope parameter for size, $\phi^{S}_{jQ}$ is the random effect of quadrat group location, and $\theta^{S}_{k}$ is the fixed parameter for the effect of the $k$th climate covariate at time $t$ ($C_{k,t}$). We include density-dependence by estimating the effect of crowding on the focal individual by other individuals of the same species. $\omega$ is the effect of crowding and $w_{t,Q}$ is the crowding experienced by the focal individual at time $t$ in quadrat $Q$.

We modeled growth as gaussian process describing genet size at time $t+1$ as a function of size at $t$ and climate covariates:

\begin{equation}
x_{ijQ,t+1} = \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{ij,t} + \omega^{G}_{j}w_{ij,t} + \theta^{G}_{jk}C_{k,t}
\end{equation}

where $x$ is log genet size and all other paramters are as described for the survival regression.

Our data allows us to track new recruits, but we cannot assign a specific parent to new genets. So, for recruitment, we work at the quadrat level and model the number of new individuals of species $j$ in quadrat $q$ recruiting at time $t+1$ as a function of quadrat "effective cover" ($A'$) in the previous year ($t$). Effective cover is a mixture of observed cover ($A$) in the focal quadrat ($q$) and the mean cover across the entire group ($Q$) of quadrats in which $q$ is located:

\begin{equation}
A'_{jq,t} = p_{j}A_{jq,t} + (1-p_{j})\bar{A}_{jQ,t}
\end{equation}

where $p$ is a mixing fraction between 0 and 1 that is estimated within the model.

We assume the number of individuals, $y$, recruiting at time $t+1$ follows a negative binomial distribution:

\begin{equation}
y_{jq,t+1} \sim \text{NegBin}(\lambda_{jq,t+1},\zeta)
\end{equation}

where $\lambda$ is the mean intensity and $\zeta$ is the size parameter. We define $\lambda$ as:

\begin{equation}
\lambda{jq,t+1} = A'_{jq,t}e^{(\gamma^{R}_{j,t} + \phi^{R}_{jQ} + \theta^{R}_{jk}C_{k,t} + \omega^{R}\sqrt{A'_{q,t}})}
\end{equation}

where $A'$ is effective cover ($\text{cm}^2$) of species $j$ in quadrat $q$ and all other terms are as in the survival and growth regressions.

#### Vital rate models: quadrat level
At the quadrat level we defined three vital rates:

1. Probability of extirpation ($S$): the probability that, for a given species, a particular quadrat will go from non-zero cover at time $t$ to zero cover at time $t+1$.
2. Cover change ($G$): the change in percent cover from time $t$ to $t+1$ for a given species within a particular quadrat.
3. Probability of colonization ($R$): the probability that, for a give species, a particular quadrat will go from zero cover at time $t$ to non-zero cover at time $t+1$.

We retain the abbreviations $S$, $G$, and $R$ from the analagous processes at the individual level. The vital rate models at the quadrat level are all based on quadrat proportional cover. Also, at the quadrat level we do not need to explicitly include a density dependent term. Since we are modeling proportional cover, we essentially get density-dependence for "free" when proportional cover is included as a covariate. That is, density-dependence emerges because quadrats with low cover gerenally increase in cover whereas quadrats with high cover generally decrease in cover.

We modeled the probability of extirpation of species $j$ in quadrat $q$ from time $t$ to $t+1$ as:

\begin{equation}
\text{logit}(S_{jq,t}) = \gamma^{S}_{j} + \phi^{S}_{jQ} + \beta^{S}_{j}x_{jq,t} + \theta^{S}_{jk}C_{k,t}
\end{equation}

where all parameters are as in Eq. 1 except that $x$ is now quadrat proportional cover. Note, however, that we do not include year random effects on the intercept $\gamma$ or the slope term $\beta$ for quadrat proportional cover. The quadrat data is inherently more sparse than the individual data from which it is aggregated, and this is especially evident when modeling rare events like extirpation and colonization. Thus, when we tried to fit random year effects, those terms did not converge.

We modeled quadrat cover change ($G$) from time $t$ to $t+1$ as:

\begin{equation}
\text{logit}(x_{jq,t+1}) = \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t}
\end{equation}

where, in this case, we do include random year effects on the intercept $\gamma$ and the slope term $\beta$. For cover change we had enough data for those terms to converge. Note that our model for quadrat cover change uses a logit transformation to link the expected cover at $t+1$ ($x_{jq,t}$) to the linear predictors. We do so because during model fitting we use a beta likelihood since the data, proportional cover, is beta distributed. The beta likelihood requires shape ($\rho$) and rate ($\eta$) parameters that can be caclulated using moment-matching:

\begin{align}
\rho &= x_{jq,t+1}\tau_{j} \\
\eta &= (1-x_{jq,t+1})\tau_{j}
\end{align}

### Model fitting
All of our analyses (model fitting and simulating) were conducted in R [@Team2013]. We used the MCMC sampler in JAGS [@Plummer2003] to estimate the posterior distributions of model parameters and the package 'r2jags' [@Su2012] to connect R to JAGS. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 50,000 iterations, after discarding an initial 50,000 iterations. We assessed convergence visually and using the @Gelman1992 diagnostic in the R package `coda' [@Plummer2006]. Scale reduction factors for all parameters were less than 1.02, indicating convergence. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each chain for all parameters to be used as randomly drawn values during population simulation.

@@

### 2014-12-19 manuscript/manuscript.tex
writing methods

Population models are important tools for predicting the impacts of
environmental change on species. But reconciling the scales at which
population models are parameterized and the scales at which
environmental changes play out remains a challenge (Freckleton et al.
2011, Queenborough et al. 2011). The major hurdle is that most
population models, at least for plant species, are built using data from
small, localized plots because parameterizing traditional population
models requires tracking the fates of individuals. These models are
difficult to scale up from the micro to meso-scales because the fitted
parameters do not fully represent the spatial variation present scales
beyond that at which the data are collected (Sæther et al. (2007)).
Thus, our ability to use population models to predict the consequences
of climate change is limited.

Recently, Freckleton et al. (2011), building on work by Hastings and
NAME, have proposed density-structured population models that focus on
the transition of populations among discrete states, rather than the
traditional approach of modeling the transitions of individuals. Such an
approach could be extremely valuable because the data needed to
parameterize density-structured population models is much easier, and
less costly, to collect (Queenborough et al. year). For example, using a
density-structured approach, one could build a population model using a
time series of annual plot-based censuses of species percent cover.
However, a major assumption of the density-structured approach is that
the aggregate dynamics of the population observed at coarse spatial
resolution faithfully represent, and correspond to, the fates of
individual plants. In other words, using a density-structured approach
requires a leap of faith that important covariates (e.g., climate
variables) at the level of the individual are captured adequately at the
population level. If we seek to forecast the impacts of climate change
on plant populations, then clearly this assumption requires testing.

@@

@@

The first step in building our single-species population models was to
fit statistical models of vital rates at both levels of inference:
individual-level and quadrat-level. Here we describe the general
statistical approach at the individual-level, but the same approach
applies at the quadrat-level. We modeled survival and growth as
functions of genet size (quadrat cover in the quadrat-level models) and
climate covariates (described in more detail below). We maintained a
consistent random effects structure for both models that included three
terms: (1) a random year effect on the intercept, (2) a random year
effect on the coefficient for plant cover (either individual or quadrat
level), and (3) a random effect of group (see Data set description) on
the intercept. Vital rates for each species are modeled separately.

We used logistic regression to model survival probability:

\begin{equation}
logit(s) = \beta_{0,t} + \beta_{s,t}x + \beta_{Q} + \beta_{c,1}\theta_{1,t}\\
+ \cdots +  \beta_{c,i}\theta_{i,t} + \varepsilon_{t}
\end{equation}

where $x$ is the log of genet size (quadrat areal cover), $\beta_{0,t}$
is a year-specific intercept, $\beta_{s,t}$ is the year-specific slope
parameter for size, $\beta_{Q}$ is the random effect of quadrat group
location, $\theta$ is a matrix of $i$ climate effects over $t$ years,
$\beta_{c,i}$ is the fixed parameter for the effect of the $i$th climate
covariate, and $\varepsilon_{t}$ is the error term. At the quadrat level
we did not fit random year effects for survival because those parameters
did not reach convergence (see ``Computing'').

We modeled growth as gaussian process describing genet size (or quadrat
cover) at time $t+1$ as a function of size at $t$ and climate
covariates:

\begin{equation}
x_{t+1} = \beta_{0,t} + \beta_{s,t}x_{t} + \beta_{Q} + \beta_{c,1}\theta_{1,t}\\
+ \cdots +  \beta_{c,i}\theta_{i,t} + \varepsilon_{t}
\end{equation}

where $x$ is genet size and all other paramters are as described for the
survival regression. For the quadrat-level approach we modeled growth as
a process describing proportional cover within a quadrat at time $t+1$
as a function of proportional cover at time $t$ and climate covariates.
Thus, instead of a gaussian process, with a normal likelihood, we
modeled growth at the quadrat level as above but with a beta likelihood
and a logit link to the linear predictors.

While our approach for modeling survival and growth is similar at the
individual and quadrat levels, the addition of new genets at the
inidividual level or occupied sites at the quadrat level requires two
separate approaches.

\subsubsection{Including climate covariates (and avoiding model
selection)}\label{including-climate-covariates-and-avoiding-model-selection}

For both the individual-level IPM (IPM) and the quadrat-based IBM (QBM)
we included four climate covariates in each vital rate regression: fall
through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and
ppt2, respectively) and mean spring temperature at \emph{t}-1 and
\emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is
the observation year. We included climate covariates as additive
effects. We did not consider interactions among climate covariates or
among climate covariates and plant size or proportional cover. Climate
effects ($\beta_{c}$s) were modeled hierarchically in each vital rate
regression so that each species specific climate effect is drawn from an
interspecific climate effect distribution. This was necessary to achieve
convergence at the quadrat level and we used the same approach at the
individual level for uniformity.

@@

All of our analyses (model fitting and simulating) were conducted in
Program R. We used the MCMC sampler in JAGS to estimate the posterior
distributions of model parameters. We obtained posterior distributions
for all model parameters from three parallel MCMC chains run for 50,000
iterations, after a 50,000 iteration burn in. We assessed convergence
visually and using the Gelman diagnostic in the R package `coda.' Scale
reduction factors for all parameters were less than 1.02, indicating
convergence. For the purposes of introducing stochasticity in our
population models, we saved the final 1,000 iterations from each chain
for all parameters to be used as randomly drawn values during population
simulation.

@@

\begin{center}\rule{3in}{0.4pt}\end{center}

@@

@@

@@

>>>>

Population models are important tools for predicting the impacts of
environmental change on species. But reconciling the scales at which
population models are parameterized and the scales at which
environmental changes play out remains a challenge (Freckleton et al.
2011, Queenborough et al. 2011). The major hurdle is that most
population models, at least for plant species, are built using data from
small, localized plots because parameterizing traditional population
models requires tracking the fates of individuals. These models are
difficult to scale up from the micro to meso-scales because the fitted
parameters do not fully represent the spatial variation present at
scales beyond that at which the data are collected (Sæther et al. 2007).
Thus, our ability to use population models to predict the consequences
of climate change is limited when we rely on individual-level data.

In contrast to individual-level data, population-level data arising from
large-scale census efforts

Recently, Freckleton et al. (2011), building on work by Taylor and
Hastings (2004), have proposed density-structured population models that
focus on the transition of populations among discrete states, rather
than the traditional approach of modeling the transitions of
individuals. Such an approach could be extremely valuable because the
data needed to parameterize density-structured population models is much
easier, and less costly, to collect.

For example, using a density-structured approach, one could build a
population model using a time series of annual plot-based censuses of
species percent cover. However, a major assumption of the
density-structured approach is that the aggregate dynamics of the
population observed at coarse spatial resolution faithfully represent,
and correspond to, the fates of individual plants. In other words, using
a density-structured approach requires a leap of faith that important
covariates (e.g., climate variables) at the level of the individual are
captured adequately at the population level. If we seek to forecast the
impacts of climate change on plant populations, then clearly this
assumption requires testing.

@@

In this paper, we model populations based on two levels of data:
individual and quadrat. The individual data is the ``raw'' data. For the
quadrat level we data we simply sum individual areal cover for each
quadrat by species. This is equivalent to a perfect census of quadrat
percent cover, so we do not need to consider measurement error. Based on
these two datasets we can compare population models built using
individual level data and aggregated quadrat level data.

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual level data we fit models for survival, growth, and
recruitment of new individuals for each species. At the quadrat level we
fit analagous models of extinction probability, percent cover
increase/decrease, and quadrat colonization for each species. We
describe the statistical models seprately since fitting the models
required different approaches at the individual and quadrat levels. All
models contain four climate covariate that we chose \emph{a priori}:
fall through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and
ppt2, respectively) and mean spring temperature at \emph{t}-1 and
\emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is
the observation year

\paragraph{Vital rate models: individual
level}\label{vital-rate-models-individual-level}

We used logistic regression to model survival probability ($S$) of genet
$i$ from species $j$ in quadrat group $Q$ from time $t$ to $t+1$:

\begin{equation}
\text{logit}(S_{ijQ,t}) = \gamma^{S}_{j,t} + \phi^{S}_{jQ} + \beta^{S}_{j,t}x_{ij,t} + \omega^{S}_{j}w_{ij,t} + \theta^{S}_{jk}C_{k,t} + \varepsilon^{S}_{t}
\end{equation}

where $x_{ij,t}$ is the log of genet size, $\gamma^{S}_{j,t}$ is a
year-specific intercept, $\beta^{S}_{j,t}$ is the year-specific slope
parameter for size, $\phi^{S}_{jQ}$ is the random effect of quadrat
group location, and $\theta^{S}_{k}$ is the fixed parameter for the
effect of the $k$th climate covariate at time $t$ ($C_{k,t}$). We
include density-dependence by estimating the effect of crowding on the
focal individual by other individuals of the same species. $\omega$ is
the effect of crowding and $w_{t,Q}$ is the crowding experienced by the
focal individual at time $t$ in quadrat $Q$.

We modeled growth as gaussian process describing genet size at time
$t+1$ as a function of size at $t$ and climate covariates:

\begin{equation}
x_{ijQ,t+1} = \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{ij,t} + \omega^{G}_{j}w_{ij,t} + \theta^{G}_{jk}C_{k,t}
\end{equation}

where $x$ is log genet size and all other paramters are as described for
the survival regression.

Our data allows us to track new recruits, but we cannot assign a
specific parent to new genets. So, for recruitment, we work at the
quadrat level and model the number of new individuals of species $j$ in
quadrat $q$ recruiting at time $t+1$ as a function of quadrat
``effective cover'' ($A'$) in the previous year ($t$). Effective cover
is a mixture of observed cover ($A$) in the focal quadrat ($q$) and the
mean cover across the entire group ($Q$) of quadrats in which $q$ is
located:

\begin{equation}
A'_{jq,t} = p_{j}A_{jq,t} + (1-p_{j})\bar{A}_{jQ,t}
\end{equation}

where $p$ is a mixing fraction between 0 and 1 that is estimated within
the model.

We assume the number of individuals, $y$, recruiting at time $t+1$
follows a negative binomial distribution:

\begin{equation}
y_{jq,t+1} \sim \text{NegBin}(\lambda_{jq,t+1},\zeta)
\end{equation}

where $\lambda$ is the mean intensity and $\zeta$ is the size parameter.
We define $\lambda$ as:

\begin{equation}
\lambda{jq,t+1} = A'_{jq,t}e^{(\gamma^{R}_{j,t} + \phi^{R}_{jQ} + \theta^{R}_{jk}C_{k,t} + \omega^{R}\sqrt{A'_{q,t}})}
\end{equation}

where $A'$ is effective cover ($\text{cm}^2$) of species $j$ in quadrat
$q$ and all other terms are as in the survival and growth regressions.

\paragraph{Vital rate models: quadrat
level}\label{vital-rate-models-quadrat-level}

At the quadrat level we defined three vital rates:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Probability of extirpation ($S$): the probability that, for a given
  species, a particular quadrat will go from non-zero cover at time $t$
  to zero cover at time $t+1$.
\item
  Cover change ($G$): the change in percent cover from time $t$ to $t+1$
  for a given species within a particular quadrat.
\item
  Probability of colonization ($R$): the probability that, for a give
  species, a particular quadrat will go from zero cover at time $t$ to
  non-zero cover at time $t+1$.
\end{enumerate}

We retain the abbreviations $S$, $G$, and $R$ from the analagous
processes at the individual level. The vital rate models at the quadrat
level are all based on quadrat proportional cover. Also, at the quadrat
level we do not need to explicitly include a density dependent term.
Since we are modeling proportional cover, we essentially get
density-dependence for ``free'' when proportional cover is included as a
covariate. That is, density-dependence emerges because quadrats with low
cover gerenally increase in cover whereas quadrats with high cover
generally decrease in cover.

We modeled the probability of extirpation of species $j$ in quadrat $q$
from time $t$ to $t+1$ as:

\begin{equation}
\text{logit}(S_{jq,t}) = \gamma^{S}_{j} + \phi^{S}_{jQ} + \beta^{S}_{j}x_{jq,t} + \theta^{S}_{jk}C_{k,t}
\end{equation}

where all parameters are as in Eq. 1 except that $x$ is now quadrat
proportional cover. Note, however, that we do not include year random
effects on the intercept $\gamma$ or the slope term $\beta$ for quadrat
proportional cover. The quadrat data is inherently more sparse than the
individual data from which it is aggregated, and this is especially
evident when modeling rare events like extirpation and colonization.
Thus, when we tried to fit random year effects, those terms did not
converge.

We modeled quadrat cover change ($G$) from time $t$ to $t+1$ as:

\begin{equation}
\text{logit}(x_{jq,t+1}) = \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t}
\end{equation}

where, in this case, we do include random year effects on the intercept
$\gamma$ and the slope term $\beta$. For cover change we had enough data
for those terms to converge. Note that our model for quadrat cover
change uses a logit transformation to link the expected cover at $t+1$
($x_{jq,t}$) to the linear predictors. We do so because during model
fitting we use a beta likelihood since the data, proportional cover, is
beta distributed. The beta likelihood requires shape ($\rho$) and rate
($\eta$) parameters that can be caclulated using moment-matching:

\begin{align}
\rho &= x_{jq,t+1}\tau_{j} \\
\eta &= (1-x_{jq,t+1})\tau_{j}
\end{align}

@@

All of our analyses (model fitting and simulating) were conducted in R
(Team 2013). We used the MCMC sampler in JAGS (Plummer 2003) to estimate
the posterior distributions of model parameters and the package `r2jags'
(Su and Yajima 2012) to connect R to JAGS. We obtained posterior
distributions for all model parameters from three parallel MCMC chains
run for 50,000 iterations, after discarding an initial 50,000
iterations. We assessed convergence visually and using the Gelman and
Rubin (1992) diagnostic in the R package `coda' (Plummer et al. 2006).
Scale reduction factors for all parameters were less than 1.02,
indicating convergence. For the purposes of introducing stochasticity in
our population models, we saved the final 1,000 iterations from each
chain for all parameters to be used as randomly drawn values during
population simulation.

@@

@@

Gelman, A., and D. B. Rubin. 1992. Inference from Iterative Simulation
Using Multiple Sequences.

Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
3rd international workshop on distributed statistical computing (dSC
2003). march.

Plummer, M., N. Best, K. Cowles, and K. Vines. 2006. CODA: Convergence
Diagnosis and Output Analysis for MCMC. R News 6:7--11.

@@

Su, Y., and M. Yajima. 2012. R2jags: A Package for Running jags from R.
http://CRAN. R-project. org/package= R2jags.

@@

Taylor, C. M., and A. Hastings. 2004. Finding optimal control strategies
for invasive species: a density-structured model for Spartina
alterniflora. Journal of Applied Ecology 41:1049--1057.

Team, R. 2013. R Development Core Team. R: A Language and Environment
for Statistical Computing.
